---
title: "L3 French Voice-Onset Time at first exposure by Spanish-English bilinguals"
subtitle: 'CASPSLaP, February 2022'
author  : "Kyle Parrish"
date    : "Rutgers University"
output:
  xaringan::moon_reader:
    chakra: "https://cdnjs.cloudflare.com/ajax/libs/remark/0.14.0/remark.min.js"
    lib_dir: libs
    css: [default, hygge, rutgers, rutgers-fonts]
    nature:
      beforeInit: ["https://www.jvcasillas.com/ru_xaringan/js/ru_xaringan.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---
```{css, echo = FALSE}
.remark-slide-content {
  font-size: 20px;
  padding: 20px 80px 20px 80px;
}
.remark-code, .remark-inline-code {
  background: #f0f0f0;
}
.remark-code {
  font-size: 24px;
}
.huge .remark-code { /*Change made here*/
  font-size: 200% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 70% !important;
}
.med .remark-code { /*Change made here*/
  font-size: 120% !important;
}
```

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE, dpi = 300)
```

```{r, load_data}
library(tidyverse)
library(sjPlot)
library(here)
library(lme4)

participants <- read.csv(here::here("data", "tidy", "participant_count.csv"))

l2_df <- read.csv(here::here("data", "tidy", "l2_subset.csv"))
l1_df <- read.csv(here::here("data", "tidy", "non_l2_subset.csv"))

# load models 
mod0 <- readRDS(here("data", "models", "full_model.RDS"))
l1_mod <- read_rds(here("data", "models", "l1_sub_model.RDS"))
l2_mod <- read_rds(here("data", "models", "l2_sub_model.RDS"))

all_desc_df <- read.csv(here("data", "tidy", "all_desc.csv"))
stim_df <- read.csv(here("data", "tidy", "stim.csv"))

total_participants <- sum(participants$l2_subset_no, participants$non_l2_no, participants$dqed_no)

total_participants_q <- sum(participants$l2_subset_no, participants$non_l2_no)

mono_df <- read.csv(here("data", "tidy", "mono_df.csv")) %>% 
  mutate(group = "monolingual") %>% 
  select(participant, text, word, duration, vot, relative_vot, language, group)


mono_participants <- unique(mono_df$participant) %>% as.data.frame()

source(here::here("scripts", "07_small_data.R"))
```


# Overview

.pull-left[
.full-width[
.content-box-red[
- How do bilinguals produce words in a new language at first exposure?

- Mexican Spanish L1-English L2 speakers produced French words.

- VOT was measured as a window into CLI.

- The production of new language (L3) words fell between L1 and L2 values. 

]
]
]

.pull-right[
```{r, echo=F}
```
]

.footnote[
(Westergaard et al., 2017)
]


---
# Introduction

.big[
- L3 learning is not L2 learning.
]

---

# Introduction

.big[
- How do L1 and L2 systems interact in L3 production?

- L3 models attempt to predict acquisition patterns when 2 languages are available

- Complicating factors: 
 - Range of L2 ultimate attainment 
 - Language use and dominance 
 - Age effects 
 
- Key questions in L3A: 

 - Does just 1 language, or do both languages affect L3A? 
 - Are bilinguals better at learning languages than monolinguals?
]

???

The difficulty of learning a new language in adulthood has a well documented history.   
Much less is known about the acquisition of a third language (L3A)

Research in third language acquisition has attempted to model the interplay between L1 and L2 language systems and their cumulative influence in the process of the acquisition of a third.

mong questions asked by third language models is whether the L1 or L2, or a combination of both languages, serves as the basis in L3 acquisition.

This question is complicated in the context of multilingualism due to the widespread diversity in bilingual populations that include wide variation in ultimate attainment in adult L2 learners, and, in the case of phonological acquisition, wide variation in the production patterns of L2 segments

---

# Research Questions

.big[
 Question 1: Does just 1 language, or do both languages affect L3A? 
 
 Question 2: Are bilinguals better at learning languages than monolinguals

L3 models are guided by these questions. 

 - What do you think? 
]

---

# Models

.big[
The **Cumulative Enhancement Model** (Flynn, Foley and Vinnitskaya, 2004)
  - Does just 1 language, or do both languages affect L3A? 
   - Both languages.
  -  Are bilinguals better at learning languages than monolinguals?
   - Yes.
]

--

- Evidence 


---

# Models
   
.big[
The **L2 Status Factor** (Bardel & Falk, 2007)
  - Does just 1 language, or do both languages affect L3A? 
   - ONLY the L2 affects the L3 in the beginning
  - Are bilinguals better at learning languages than monolinguals?
   - No.
]

--

- Evidence 

---

# Models   
  
.big[
The **Typological Primacy Model** (Rothman, 2011)
  - Does just 1 language, or do both languages affect L3A? 
   - Only 1 language affects the L3 - closest typological match. 
  - Are bilinguals better at learning languages than monolinguals?
   - No.
]

--

- Evidence 

---

# Models

.big[
The **Linguistic Proximity Model** (Westergaard et al., 2017)
  - Does just 1 language, or do both languages affect L3A? 
   - Both
  - Are bilinguals better at learning languages than monolinguals?
   - Maybe.
]   

--

- Evidence 

 
---

# Literature Review 

**Broad questions** 
 - How do language systems interact during L3 acquisition?
 
--

  - Possible sources of **progressive transfer:** 
  - L1 alone 
  - L2 alone
  - combination of both languages

--

**Models**
- L3 models: predictions vary. 
 - TPM - closest typological language 
 - L2 Status Factor - L2 by default influences the L3.
 - LPM - either or both language may affect L3 productions.
 
--

**Previous Work**
- Work in L3 VOT production mixed:
  - L2 influence
  - Hybrid values 
  
---

# Literature Review 

**Previous Work (cont)**

- A formal model of L3 phonological acquisition has not been proposed 
  - The expansion of L2 Speech models have been proposed 
    - SLM (Flege & Bohn, 2021)
    - PAM (Best & Tyler, 2007)
    - L2LP (van Leussen & Escudero, 2015)
 
- Cross-linguistic similarity of segments is predicted to determine the learnability of segments in a second language. 

--

- In the case of L3 learners, two phonological systems exist to which the L3 segments could be compared.

- It is not well known how assimilation of L3 sounds occur with L1 and L2 categories.

---

# Literature Review 

### How can the body of work to date be contributed to? 

.pull-left[
- Low sample sizes in L3 work 
]

--

.pull-left[
- Few studies examine first exposure to the L3 
]

--

.pull-left[
- Proposal: **higher sample size of bilinguals at first exposure**
]
---

# Literature Review

.big[
- Cross-linguistic VOT
 - **Spanish** and **French** - true-voicing languages 
 - **English** - aspirating language 
]

---

# Methods 

.pull-left[.big[
- **RQ:** When Spanish-English bilinguals produce French words at first exposure, will their VOT productions be more L1 or L2 like?  
]]

--

.pull-right[.big[
- **Predictions:** It is possible that hybrid values reported in the literature are due to sampling error or phonological acquisition - it was predicted that **L3 VOT would be practically equivalent to L2 VOT.**
]]

---
# Materials 

**VOT measurements**
- French shadowing task 

- English elicited Production task

- Spanish elicited Production task 

--

**Language proficiency and background**
- Self-rating on a 1-7 scale

- Language Questionnaire 

---

# Materials 

.pull-left[
- **Stimuli**
 - Word initial stops of either one or two syllables, with stress on the initial syllable. 
 
 
- **Word lists**
 - *Spanish:* tiro, tema, talla, quiso, queja, cama, piso, pena, pato
 - *English:* tipping, teller, tacky, penny , pass, parrot, kitten, kennel, cabbage
 - *French:* tir, terre, tasse, quitte, quelle, pile, pere, patte
]

--

.pull-right[
- **Tokens for shadowing task**
 - Produced by a native French speaker
 - Mean relative VOT = `r mean(stim_df$relative_vot) %>% round(digits = 3)`, sd = `r sd(stim_df$relative_vot) %>% round(digits = 3)`
]

---

# Procedure 

- The elicited production task and shadowing task were programmed in Labvanced, a browser based software designed for online research. 

- All tasks were given in one session.

--

- Order of the languages was counterbalanced and the stimulus order was randomized. 

- Words produced in isolation

---

# Participants 

- A total of `r total_participants` participants took part in the experiment.   

- The data of `r participants$dqed_no` participants were removed from the final data set, since these bilinguals did not produce distinct Spanish-English VOT values based on a t-test. 

- Thus, the data of a total of **`r total_participants_q`** participants were included in the final analysis. 


---

# Participants 

- Participants were **L1 Mexican Spanish**, **L2 English** late bilingual speakers recruited from the online platform prolific.co 

- **Location:** All participants were born and lived in Mexico at the time of the study.

--

- **Use:** Assumed to be primarily Spanish

- **Proficiency:** Self-rated + phonetic corroboration

--

- **Dominance:** Filtering location was intended to find Spanish-dominant speakers

--

- All participants reported not speaking or having studied a third language. 

--

```{r}
filt_bio <- bio_data %>% 
  filter(participant %in% all_df$participant)

data.frame("Factor" = c("AoA", "Current age", "Proficiency"),
           "Mean" = c(mean(filt_bio$age), mean(filt_bio$current_age), mean(filt_bio$self_prof)),
           "SD" = c(sd(filt_bio$age), sd(filt_bio$current_age), sd(filt_bio$self_prof))) %>% 
  knitr::kable(digits = 2)

```

---

# Data segmentation procedure 

- Productions were force aligned using WebMaus and hand-corrected using a PRAAT script. 

- VOT was marked at the closest zero-crossing to the onset of visible aspiration and the zero-crossing before before the onset of the vowel.


---

# Statistical Analysis 

- All analyses were carried out in `R`

- **Sample size justification**: Power analysis

--

- **Inclusion and subsetting**: T.tests 

--

- **Quantify the effect of language**: GLMM

--

- **Determine equivalence**: TOST

---
# Results - all participants 

```{r}
all_df <- rbind(l1_df, l2_df)
all_part <- length(unique(all_df$participant))
```

.pull-left[.tiny[
n = `r all_part`

```{r, fig.height=4, fig.width=5}
all_df %>% 
  ggplot(aes(x = relative_vot, y = language)) + geom_boxplot() + xlim(0,.3)
```

```{r}
all_df %>% 
  group_by(language) %>% 
  summarise(`Relative VOT` = mean(relative_vot), `SD` = sd(relative_vot)) %>% 
  knitr::kable(digits = 3)
```
]]

.tiny[.pull-right[
```{r}
tab_model(mod0)
```
]]

---

# Averages per segment full dataset

```{r}
all_df %>% 
  group_by(language, text) %>% 
  summarise(`Relative VOT` = mean(relative_vot), `SD` = sd(relative_vot)) %>% 
  filter(text == "k") %>% 
  knitr::kable(digits = 3)
```

```{r}
all_df %>% 
  group_by(language, text) %>% 
  summarise(`Relative VOT` = mean(relative_vot), `SD` = sd(relative_vot)) %>% 
  filter(text == "t") %>% 
  knitr::kable(digits = 3)
```

```{r}
all_df %>% 
  group_by(language, text) %>% 
  summarise(`Relative VOT` = mean(relative_vot), `SD` = sd(relative_vot)) %>% 
  filter(text == "p") %>% 
  knitr::kable(digits = 3)
```

---

# Results: L2 subset

.pull-left[.tiny[

Subsetting procedure: inconclusive French-English t.tests 

**Hybrid values**  n = `r participants$l2_subset_no`

```{r fig1, fig.height=4, fig.width=5, dpi=300}
l2_df %>% 
  ggplot(aes(x = relative_vot, y = language)) + geom_boxplot() + xlim(0,.3)
```

```{r}
l2_df %>% 
  group_by(language) %>% 
  summarise(`Relative VOT` = mean(relative_vot), `SD` = sd(relative_vot)) %>% 
  knitr::kable(digits = 4)
```
]]

.tiny[.pull-right[
```{r, dpi=300}
tab_model(l2_mod)
```
]]

---

# L2 subset - Test of Equivalence 

.pull-left[
```{r, fig.height = 8.5, fig.width = 8}
l2_subset_df <- l2_df
l2_subset_df_fr <- l2_subset_df %>% 
  filter(language == "french")
l2_subset_df_en <- l2_subset_df %>% 
  filter(language == "english")

# TOST to see whether a subset of participants are practically equivalent with
# eng and french 

TOSTER::TOSTtwo(n1 = length(unique(l2_subset_df$participant)),
                n2 = length(unique(l2_subset_df$participant)), 
                   m1 = mean(l2_subset_df_fr$relative_vot), 
                   m2 = mean(l2_subset_df_en$relative_vot), 
                   sd1 = sd(l2_subset_df_fr$relative_vot), 
                   sd2 = sd(l2_subset_df_en$relative_vot),
                   low_eqbound_d = -.4, 
                   high_eqbound_d = .4)
```
]

---

# Individual differences - L2 subset

```{r, fig.height=3, fig.width=6, dpi=300}
tidy_df <- read.csv(here("data", "tidy", "tidy_df_2.csv"))

tidy_df$participant <- as.factor(tidy_df$participant)
 
temp_4 <- tidy_df %>% 
  filter(fr_eng_p > .05) %>%  
  filter(sp_eng_p < .05)

l2_subset <- unique(temp_4$participant)

# L2 subset 
tidy_df %>% 
  filter(participant %in% l2_subset) %>% 
  group_by(participant, language) %>%
  summarise(vot = mean(relative_vot)) %>% 
  pivot_wider(names_from = language, values_from = vot) %>% 
  mutate("English French difference" = english - french,
         "English Spanish difference" = english - spanish) %>%  
  select(participant, `English French difference`, `English Spanish difference`) %>% 
  pivot_longer(c(`English French difference`, `English Spanish difference`), names_to = "pair", values_to = "difference") %>% 
  ggplot(aes(x = difference, y = participant, color = pair)) + geom_point() +
  ggtitle("L2 subset") + geom_vline(xintercept = 0) +
  geom_vline(xintercept = sd(tidy_df$relative_vot)*.4, linetype = "dashed") +
  geom_vline(xintercept = sd(tidy_df$relative_vot)*-.4, linetype = "dashed")
```

---

# Results: L1 subset

.pull-left[.tiny[

Subsetting procedure: p < .05 French-English t.tests 

**Heavy L1 influence** n = `r participants$non_l2_no`

```{r fig2, fig.height = 4, fig.width = 5, dpi=300}
l1_df %>% 
  ggplot(aes(x = relative_vot, y = language)) + geom_boxplot() + xlim(0,.3)
```

```{r}
l1_df %>% 
  group_by(language) %>% 
  summarise(`Relative VOT` = mean(relative_vot), `SD` = sd(relative_vot)) %>% 
  knitr::kable(digits = 3)
```
]]

.tiny[.pull-right[
```{r, dpi=300}
tab_model(l1_mod)
```
]]
---

# L1 subset - Test of Equivalence 

.pull-left[
```{r, fig.height = 8.5, fig.width = 8}
non_l2_subset_df_fr <- l1_df %>% 
  filter(language == "french")
non_l2_subset_df_sp <- l1_df %>% 
  filter(language == "spanish")

TOSTER::TOSTtwo(n1 = length(unique(l1_df$participant)),
                   n2 = length(unique(l1_df$participant)), 
                   m1 = mean(non_l2_subset_df_fr$relative_vot), 
                   m2 = mean(non_l2_subset_df_sp$relative_vot), 
                   sd1 = sd(non_l2_subset_df_fr$relative_vot), 
                   sd2 = sd(non_l2_subset_df_sp$relative_vot),
                   low_eqbound_d = -.4, 
                   high_eqbound_d = .4)

```
]


---
# Individual Differences - L1 subset

```{r, fig.height=3, fig.width=6, dpi=300}
# L1 subset 
tidy_df %>% 
  filter(!participant %in% l2_subset) %>% 
  group_by(participant, language) %>%
  summarise(vot = mean(relative_vot)) %>% 
  pivot_wider(names_from = language, values_from = vot) %>% 
  mutate("sp_fr_diff" = french - spanish,
         "sp_eng_diff" = english - spanish) %>%  
  select(participant, sp_eng_diff, sp_fr_diff) %>% 
  pivot_longer(c(`sp_eng_diff`, `sp_fr_diff`), names_to = "pair", values_to = "difference") %>% 
  ggplot(aes(x = difference, y = participant, color = pair)) + geom_jitter() +
  ggtitle("L1 subset") + geom_vline(xintercept = 0) +
  geom_vline(xintercept = sd(tidy_df$relative_vot)*.4, linetype = "dashed") +
  geom_vline(xintercept = sd(tidy_df$relative_vot)*-.4, linetype = "dashed")
```

---

# Post-hoc analyses 

```{r}
all_desc_df %>% 
  knitr::kable(digits = 3, caption = "Effect sizes of each subset per language pairing")

```
---

# Post-hoc analyses 

```{r, fig.height = 3, fig.width = 7}
all_desc_df %>% 
  ggplot(aes(x = eff_size, y = set, color = pair)) + geom_point() +
  geom_linerange(aes(xmin = ci_low, xmax = ci_hi)) +
  geom_vline(xintercept = 0) + geom_vline(xintercept = .4, linetype = "dashed", alpha = .5) +
  geom_vline(xintercept = -.4, linetype = "dashed", alpha = .5) +
  ggtitle("Effect sizes per dataset and language pairing")
```


---

# Limitations 

.big[
- A rich background questionnaire was not taken.

- Subset results may be sampling error. 

- These individual differences could be explored in future research.
]

---

# Discussion and Conclusions 

- Participants are primarily influenced by one of their source languages but to different degrees

--

 - L1 influence, L3 VOT ~ L1 values, 
 - L2 influence, L3 VOT values fell between L1 and L2 values
 
--

```{r}
power_df %>% 
  select(power, subset, pairing) %>% 
  knitr::kable(caption = "Post-hoc power anlyses")

```

---

# Discussion and Conclusions  

- Statistical power > .8 

- First exposure bilinguals allow for the teasing apart of facilitative influence of a previously known language and phonological acquisition.

--

- **Future project:** 
 - Perception task - PAM phoneme categorization task.
 - Same group, but an aspirated L2, such as German.  


---

# Future directions and questions 

- If the L3 stimulus were more L2-like 

--

- Detailed case studies
 
---

# Future directions and questions 

- Phonetics or language influence? 


.pull-left[
```{r, fig.height=4, fig.width=5}
l2_df %>% 
  ggplot(aes(x = relative_vot, y = language)) + geom_boxplot() + xlim(0,.3) + ggtitle("L2 subset")
```
]

.pull-right[
```{r, fig.height=4, fig.width=5}
l1_df %>% 
  ggplot(aes(x = relative_vot, y = language)) + geom_boxplot() + xlim(0,.3) + ggtitle("L1 subset")
```
]


